Reanalysis Step 8: Patient-Level Table 1

### 2025-09-05 reanalysis-8-table1.qmd
table1

---
title: "Reanalysis Step 7: Sample Comparison Table"
subtitle: "Comparing Original, Complete Cases, and Imputed Data"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: html
---

@@

# Initialize logging
log_file <- init_log("sample_comparison")
```

@@

log_message(paste0("Original data after exclusions: ", nrow(original_complete), " observations"))

@@

@@

log_message(paste0("Submitted (included) data: ", nrow(submitted_data), " observations"))
log_message(paste0("Excluded data: ", nrow(excluded_data), " observations"))

@@

log_message(paste0("Loaded ", n_imputations, " imputed datasets"))
```

@@

# Sample sizes
sample_sizes <- data.frame(
  Dataset = c("Original (after exclusions)", "Submitted (included)", "Excluded", "Imputed (each)"),
  N = c(nrow(original_complete), nrow(submitted_data), nrow(excluded_data), nrow(imputed_datasets[[1]]))
)

@@

**Note:** The "Excluded" column shows patients who were excluded from the original submitted analysis because they lacked 2-year outcome data. These patients have substantial missing data, which is why many cells show 0 (NaN%) - indicating all values are missing for those variables. The imputation process fills in these missing values, allowing us to include these patients in the analysis.

>>>>

---
title: "Reanalysis Step 8: Patient-Level Table 1"
subtitle: "Comparing Original, Complete Cases, and Imputed Data at Patient Level (Best Eye)"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: html
---

@@

# Initialize logging
log_file <- init_log("patient_level_table1")
```

@@

# Filter to patient level - best eye (lowest outcome_VA_logMAR)
original_complete <- original_complete |>
  arrange(outcome_VA_logMAR) |>
  group_by(e_mrn_deidentified) |>
  slice(1) |>
  ungroup()

log_message(paste0("Original data after exclusions (patient-level): ", nrow(original_complete), " patients"))

@@

# Filter to patient level - best eye (lowest outcome_VA_logMAR)
imputation_data <- imputation_data |>
  arrange(outcome_VA_logMAR) |>
  group_by(e_mrn_deidentified) |>
  slice(1) |>
  ungroup()

@@

log_message(paste0("Submitted (included) data: ", nrow(submitted_data), " patients"))
log_message(paste0("Excluded data: ", nrow(excluded_data), " patients"))

@@

# Filter each imputed dataset to patient level - best eye
for (i in 1:n_imputations) {
  imputed_datasets[[i]] <- imputed_datasets[[i]] |>
    arrange(outcome_VA_logMAR) |>
    group_by(e_mrn_deidentified) |>
    slice(1) |>
    ungroup()
}

log_message(paste0("Loaded ", n_imputations, " imputed datasets (patient-level)"))
```

@@

# Sample sizes
sample_sizes <- data.frame(
  Dataset = c("Original (after exclusions, patient-level)", "Submitted (included, patient-level)", "Excluded (patient-level)", "Imputed (each, patient-level)"),
  N = c(nrow(original_complete), nrow(submitted_data), nrow(excluded_data), nrow(imputed_datasets[[1]]))
)

@@

**Note:** 
- All analyses are at the patient level, selecting the best eye (lowest outcome_VA_logMAR) for each patient.
- The "Excluded" column shows patients who were excluded from the original submitted analysis because they lacked 2-year outcome data. These patients have substantial missing data, which is why many cells show 0 (NaN%) - indicating all values are missing for those variables. 
- The imputation process fills in these missing values, allowing us to include these patients in the analysis.
